The unphysicality of Hilbert spaces

### 2023-07-26 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Fixed minor issues in the bibliography and resolved minor issues in the appendix

For example, suppose we want to define volumes in three-dimensional Euclidean space $\mathbb{R}^3$. Formally, we want to find a measure $\mu(U)$ that returns the volume of a region $U$. Naively, we could take a region to be any subset of $\mathbb{R}^3$. However, this leads to the Banach–Tarski paradox,\cite{tao2021introduction} which states that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls each of the same volume as the original. Clearly, this does not make physical sense. However, we get a consistent definition if we restrict regions to be Borel sets, instead of arbitrary subsets. Therefore, bigger is not always better in mathematics:\footnote{In another work\cite{AoPBook} we show that the Borel sets are exactly those sets that can be associated with experimental procedures. We have often found that physical requirements naturally map to well-behaved spaces.} we need to look more closely at what the math dragged in.

@@

%TODO reference formatting for 29

@@

\begin{prop}[Inner product is an entropic structure]\label{prop_innerProductIsEntropy}
Let $\mathcal{H}$ be a complex inner product space. Let $I : \mathcal{H} \times \mathcal{H} \to \mathbb{R}$ be the von Neumann entropy of the equal mixture of two pure states. That is
\begin{equation}
I(\psi_1, \psi_2) = S\left(\frac{1}{2}|\psi_1\>\<\psi_1| + \frac{1}{2}|\psi_2\>\<\psi_2|\right).
\end{equation}
Let $p : \mathcal{H} \times \mathcal{H} \to \mathbb{R}$ be the square of the inner product. That is
\begin{equation}
p(\psi_1, \psi_2) = |\<\psi_1| \psi_2\>|^2.
\end{equation}
Then $p$ and $I$ can be reconstructed from each other. That is, there exists an invertible function $f$ such that
\begin{equation}
I(\psi_1, \psi_2) = f(p(\psi_1, \psi_2)).
\end{equation}
\end{prop}

@@

As $\rho$ is in a diagonal form, the entropy is given by
\begin{equation}\label{entropy}
	H(\rho) = H\left(\frac{1+|\langle\psi|\phi\rangle|}{2}, \frac{1-|\langle\psi|\phi\rangle|}{2}\right).
\end{equation}
If we define $p(\psi, \phi) = |\langle\psi|\phi\rangle|$ and $I(\psi, \phi) = H( \frac{1}{2}|\psi\>\<\psi| + \frac{1}{2}|\phi\>\<\phi|)$ we have %TODO: Define f. It looks like a von Neumann entropy, but we haven't used the notation f(p) before.
\begin{align}\label{entropy}
	f(p) & = - \frac{1+p}{2} \log \frac{1+p}{2} 
	- \frac{1-p}{2} \log \frac{1-p}{2} \\
	I(\psi, \phi) &= f(p(\psi, \phi)).
\end{align}
Note that $f$ restricted to the domain $[0,1]$, the domain on which probability is defined, is invertible.
\end{proof}

@@

A naive idea would be to require that the expectation value for all possible variables converges. Unfortunately, this doesn't work. Suppose we have a distribution $\rho(x)$ over position $x$. We would require the expectation of any function to be finite, which would include $e^x$, $e^{e^x}$, $e^{e^{e^x}}$, and so on. The only way this can work is if the function has compact support. Physically, this seems too restrictive as it excludes functions like normal distributions. At a closer investigation, the Paley-Wiener theorem\cite{fourierCompact}\footnote{%TODO: Delete this footnote?
Maybe Prop. 3.115 in \cite{moretti_spectral_2017}?} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\cite{markushevich2014entire} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform. Therefore the space of functions with compact support on phase space gives us finite expectation value for all functions of positions and momentum.}

>>>>

For example, suppose we want to define volumes in three-dimensional Euclidean space $\mathbb{R}^3$. Formally, we want to find a measure $\mu(U)$ that returns the volume of a region $U$. Naively, we could take a region to be any subset of $\mathbb{R}^3$. However, this leads to the Banach–Tarski paradox,\cite{tao2021introduction} which states that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls each of the same volume as the original. Clearly, this does not make physical sense. However, we get a consistent definition if we restrict regions to be Borel sets, instead of arbitrary subsets. Therefore, bigger is not always better in mathematics:\footnote{In another work\cite{aop-book} we show that the Borel sets are exactly those sets that can be associated with experimental procedures. We have often found that physical requirements naturally map to well-behaved spaces.} we need to look more closely at what the math dragged in.

@@

@@

\begin{prop}[Inner product is an entropic structure]\label{prop_innerProductIsEntropy}
Let $\mathcal{H}$ be a complex inner product space. Let $I : \mathcal{H} \times \mathcal{H} \to \mathbb{R}$ be the von Neumann entropy of the equal mixture of two pure states. That is
\begin{equation}
I(\psi, \phi) = S\left(\frac{1}{2}|\psi\>\<\psi| + \frac{1}{2}|\phi\>\<\phi|\right).
\end{equation}
Let $p : \mathcal{H} \times \mathcal{H} \to \mathbb{R}$ be the square of the inner product. That is
\begin{equation}
p(\psi, \phi) = |\<\psi| \phi\>|^2.
\end{equation}
Then $p$ and $I$ can be reconstructed from each other. That is, there exists an invertible function $f$ such that
\begin{equation}
I(\psi, \phi) = f(p(\psi, \phi)).
\end{equation}
\end{prop}

@@

As $\rho$ is in a diagonal form, the entropy is given by
\begin{equation}\label{entropy}
	H(\rho) = H\left(\frac{1+|\langle\psi|\phi\rangle|}{2}, \frac{1-|\langle\psi|\phi\rangle|}{2}\right).
\end{equation}
If we define $p(\psi, \phi) = |\langle\psi|\phi\rangle|$ and $I(\psi, \phi) = H( \frac{1}{2}|\psi\>\<\psi| + \frac{1}{2}|\phi\>\<\phi|)$ we have
\begin{align}\label{entropy}
	I(\psi, \phi) &= f(p(\psi, \phi))
\end{align}
where
\begin{align}\label{entropy}
	f(p) & = - \frac{1+p}{2} \log \frac{1+p}{2} 
	- \frac{1-p}{2} \log \frac{1-p}{2}.
\end{align}
Note that $f$ restricted to the domain $[0,1]$, the domain on which probability is defined, is invertible. We have thus proved the proposition.
\end{proof}

@@

A naive idea would be to require that the expectation value for all possible variables converges. Unfortunately, this doesn't work. Suppose we have a distribution $\rho(x)$ over position $x$. We would require the expectation of any function to be finite, which would include $e^x$, $e^{e^x}$, $e^{e^{e^x}}$, and so on. The only way this can work is if the function has compact support. Physically, this seems too restrictive as it excludes functions like normal distributions. At a closer investigation, the Paley-Wiener theorem\cite{fourierCompact} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\cite{markushevich2014entire} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform. Therefore the space of functions with compact support on phase space gives us finite expectation value for all functions of positions and momentum.}
